```{r}
setwd("~/Desktop/FXR/Fxr/Data")
fxr_data = na.omit(read.csv("gene count.csv", header = TRUE, stringsAsFactors = FALSE, row.names = 1))[,-c(11)]
head(fxr_data)
```

```{r}
str(fxr_data$X)
```

```{r}
row.names(fxr_data) = fxr_data$X
fxr_data = fxr_data[,-1]
head(fxr_data)
```



```{r}
setwd("~/Desktop/FXR/Fxr/Data")
design_file = read.csv("Liver_mRNA_Seq_sampleID.csv", stringsAsFactors = FALSE, row.names = 1)
head(design_file)
```

```{r}
metadata = design_file[-c(11),c(4:6)]
head(metadata)
```

```{r}
library(DESeq2)
library(dplyr)
library(ggplot2)
library(hexbin)
library(readr)
```


```{r}
dds = DESeqDataSetFromMatrix(countData = fxr_data, colData = metadata, design = ~Diet)
```

```{r}
dds = dds[rowSums(counts(dds))>10,]
```

```{r}
vsd = vst(dds, blind = TRUE)
vsd_mat = assay(vsd)
vsd_cor = cor(vsd_mat)
library(pheatmap)
pheatmap(vsd_cor, annotation = select(metadata, c(Sex, Phenotype, Diet)))
```

```{r}
plotPCA(vsd, intgroup = c("Phenotype", "Diet", "Sex")) + geom_label(aes(label = row.names(metadata)))
```

```{r}
only_HFD = fxr_data[, c(1:10)]
meta_HFD = metadata[c(1:10),]
```


```{r}
dds_control = DESeqDataSetFromMatrix(countData = fxr_data[,c(1:10)], colData = metadata[c(1:10),], design = ~Phenotype)
```

```{r}
vsd_control = vst(dds_control, blind = TRUE)
vsd_mat_control = assay(vsd_control)
vsd_cor_control = cor(vsd_mat_control)
```

```{r}

plotPCA(vsd_control, intgroup = c("Phenotype", "Sex"))

```

```{r}
dde_control = DESeq(dds_control)
res_control = results(dde_control, contrast = c("Phenotype", "Liver KO", "WT"))
res_control = lfcShrink(dde_control, contrast = c("Phenotype", "Liver KO", "WT"), res = res_control)
```

extract the data and subset for differentially expressed genes

```{r}
res_control = data.frame(res_control)
res_control = subset(res_control, padj < 0.05)
```

Write data to file
```{r}
setwd("~/Desktop/FXR/Fxr/Data")
write.csv(res_control, "Liver KO vs WT HFD.csv")
```




```{r}
only_XN = fxr_data[, c(11:19)]
meta_XN = metadata[c(11:19),]
```

```{r}
dds_tx = DESeqDataSetFromMatrix(countData = fxr_data[,c(11:19)], colData = metadata[c(11:19),], design = ~Phenotype)
```
```{r}
vsd_tx = vst(dds_tx, blind = TRUE)
vsd_mat_tx = assay(vsd_tx)
vsd_cor_tx = cor(vsd_mat_tx)
```

```{r}

plotPCA(vsd_tx, intgroup = c("Phenotype", "Sex")) #+ geom_label(aes(label = row.names(meta_XN)))

```



```{r}
dde_tx = DESeq(dds_tx)
res_tx = results(dde_tx, contrast = c("Phenotype", "Liver KO", "WT"))
res_tx = lfcShrink(dde_tx, contrast = c("Phenotype", "Liver KO", "WT"), res = res_tx)
```

extract the data and subset for differentially expressed genes

```{r}
res_tx = data.frame(res_tx)
res_tx = subset(res_tx, padj < 0.05)
```

Write data to file
```{r}
setwd("~/Desktop/FXR/Fxr/Data")
write.csv(res_tx, "Liver KO vs WT XN.csv")
```

Function to get Symbol from ensemblID
Parameter:
  data: dataframe with gene symbol as row.names.
  keys: vector of symbols
```{r}
ensemblID_symbol = function(data, keys){
  require(org.Mm.eg.db)
  require(tidyr)
  hash.table = drop_na(AnnotationDbi::select(org.Mm.eg.db,
                                   keys = keys,
                                   keytype = "ENSEMBL",
                                   columns = "SYMBOL"))
  #hash.table = drop_na(hash.table)
  data_subset = data[hash.table$ENSEMBL, ]
  data_subset$Symbol = as.character(hash.table$SYMBOL)
  
  
  return(data_subset)
}
```

```{r}
setwd("~/Desktop/FXR/Fxr/Data")
XN = read.csv("Liver KO vs WT XN.csv", stringsAsFactors = FALSE, row.names = 1)
```

```{r}
emsemble = row.names(XN)
```

```{r}
new_data = ensemblID_symbol(XN, emsemble)
```
```{r}
length(unique(new_data$Symbol))
```

```{r}
ab = table(new_data$Symbol)
ab[ab>1]
```


```{r}
 setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
write.csv(new_data, "Liver KO vs WT XN.csv")
```


```{r}
setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
HFD = read.csv("Liver KO vs WT HFD.csv", stringsAsFactors = FALSE)[,c(1,3)]
XN = read.csv("Liver KO vs WT XN.csv", stringsAsFactors = FALSE)[,c(1,3)]
```

```{r}
colnames(HFD) = c("Symbol", "log2FC HFD")
colnames(XN) = c("Symbol", "log2FC XN")
```

```{r}
HFD_XN = merge(HFD, XN, by = "Symbol", all = TRUE)
```

```{r}
row.names(HFD_XN) = HFD_XN$Symbol
```


```{r}
ensemble_symbol = function(data, keys){
  require(org.Mm.eg.db)
  require(tidyr)
  hash.table = drop_na(AnnotationDbi::select(org.Mm.eg.db,
                                   keys = keys,
                                   keytype = "ENSEMBL",
                                   columns = "SYMBOL"))
  hash.table = drop_na(hash.table)
  data_subset = data[hash.table$ENSEMBL, ]
  data_subset$EntrezID = as.character(hash.table$SYMBOL)
  data_subset$Symbol = row.names(data_subset)
  row.names(data_subset) = data_subset$SYMBOL
  
  
  return(data_subset)
}
```

```{r}
HFD_XN = ensemble_symbol(HFD_XN, keys = row.names(HFD_XN))
```


```{r}
setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
write.csv(HFD_XN, "merged de genes HFD and HFD_XN.csv")
```

```{r}
length(unique(c(HFD$Symbol, XN$Symbol)))
```

```{r}
setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
data = read.csv("merged de genes HFD and HFD_XN.csv", stringsAsFactors = FALSE, row.names = 1)
head(data)
```

```{r}
HFD_XN_only = data[(is.na(data$log2FC.HFD) & !(is.na(data$log2FC.XN))),]
```

```{r}
setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
write.csv(HFD_XN_only, "unique for HFD XN.csv")
```

```{r}
HFD_only = data[((!is.na(data$log2FC.HFD)) & (is.na(data$log2FC.XN))),]
```

```{r}
setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
write.csv(HFD_only, "unique for HFD.csv")
```

```{r}
shared = data[((!is.na(data$log2FC.HFD)) & !(is.na(data$log2FC.XN))),]
```

```{r}
setwd("C:/Users/trant/Desktop/FXR/Fxr/Data")
write.csv(shared, "shared between HFD and HFD+XN.csv")
```

```{r}
only_WT = fxr_data[, c(1:10)]
meta_HFD = metadata[c(1:10),]
```


```{r}
dds = DESeqDataSetFromMatrix(countData = fxr_data[,c(1:6, 11:14)], colData = metadata[c(1:6, 11:14),], design = ~Diet)
```


```{r}
vsd = vst(dds, blind = TRUE)
vsd_mat = assay(vsd)
vsd_cor = cor(vsd_mat)
```

```{r}

plotPCA(vsd, intgroup = "Diet")

```

```{r}
dde = DESeq(dds)
res = results(dde, contrast = c("Diet", "HFD_XN", "HFD"))
res = lfcShrink(dde, contrast = c("Diet", "HFD_XN", "HFD"), res = res)
```

extract the data and subset for differentially expressed genes

```{r}
res = data.frame(res)
res = subset(res, padj < 0.05)
```

Write data to file
```{r}
setwd("~/Desktop/FXR analysis/Fxr/Data")
write.csv(res, "WT XN vs HFD.csv")
```

```{r}
dds = DESeqDataSetFromMatrix(countData = fxr_data[c(7:10, 15:19)], colData = metadata[c(7:10, 15:19),], design = ~Diet)
```

```{r}
vsd = vst(dds, blind = TRUE)
vsd_mat = assay(vsd)
vsd_cor = cor(vsd_mat)
```

```{r}

plotPCA(vsd, intgroup = "Diet")

```

```{r}
dde = DESeq(dds)
res = results(dde, contrast = c("Diet", "HFD_XN", "HFD"))
res = lfcShrink(dde, contrast = c("Diet", "HFD_XN", "HFD"), res = res)
```

extract the data and subset for differentially expressed genes

```{r}
res = data.frame(res)
res = subset(res, padj < 0.05)
```

Write data to file
```{r}
setwd("~/Desktop/FXR analysis/Fxr/Data")
write.csv(res, "Liver KO XN vs HFD.csv")
```


